A re-examination of the role of friction in the original Social Force Model

### 2019-06-22 paper.tex
ordenamiento de citas

\textcolor{red}{In the last decades, many microscopic models 
for crowd dynamics were developed. The class of force-based model was proposed by Hirai 
and Tarui (1975) introducing the class of \textit{force based} models 
\cite{Hirai}. These models represent the interaction between 
pedestrians and their environment using different kinds of forces inspired in physical systems. For 
example, Okazaki (1979) modeled the movement of pedestrians as 
a magnetized object immersed a magnetic in field \cite{Okazaki}.}\\

@@

\textcolor{red}{ Our work focuses on data from Ref.~\cite{helbing3}. 
Further measurements can be found in Refs.~\cite{Boltes,Dridi1,Dridi2,Baqui1,lohner1}.
The PedFlow model developed by L\"ohner \textit{et al}. Ref.~\cite{lohner2}  
is an alternative to the Helbing model. Togashi \textit{et al}.  Ref.~\cite{Togashi1} 
improved the calibration of this model using a Kalman filter ensemble. }\\

The high density regime appears to be the most cumbersome one. Caution was 
claimed when (automatically) transferring the usual ``calibrated'' parameters 
of the SFM to this regime. It was argued that the pedestrians' body size 
distribution and the ``situational context'' are somewhat responsible for the 
unexpected departure from these parameters \cite{johansson1,kwak}. But other 
researchers pointed out that this departure actually expresses the lack of a 
mechanism to properly handle the pedestrians' ``required space to move''. Some 
modifications to the basic SFM were then proposed to overcome this difficulty 
\cite{parisi2,seyfried2}. \\

@@

Inspired from vehicular traffic dynamic studies, many researches on pedestrian dynamics focus their attention on the relation between the flow and the density of a moving crowd. This relation is represented by ``fundamental diagram" and it has become one of the most common ways to characterize the pedestrians' dynamics along a corridor in unidirectional and bidirectional flows \cite{fruin1,mori1,polus1,seyfried1,jelic1}. \\

@@

We implemented special modules in C++ for upgrading the LAMMPS 
capabilities to attain the Social Force Model simulations. We also checked over 
the LAMMPS output with previous computations (see Refs. \cite{Dorso1, 
Dorso2,Dorso3, Dorso4,Dorso6}).\\

@@

%\begin{align*} 
%v(\rho_{max}) &< \frac{5v_d}{\rho_{max}} \leq \frac{5}{9} v_d\\
%\end{align*}

@@

Recall from Eq.(\ref{granular}), that the friction force depends on the compression and the relative velocity among pedestrians. The compression levels remain the same in the three cases since the compression only depends on the density (which is fix at $\rho=6$ for the three color maps). Thus, the differences between Figs.~\ref{abswg_width10}, \ref{abswg_width15} and \ref{abswg_width22} can only explained by the increment of the relative velocity between individuals. \\

@@

A boundary condition needs to be imposed in order to solve 
Eqs.~(\ref{eqn_appendix_2:3}) for $\alpha\neq 0$. We fix 
$v_i=v_{i+1}$ at the middle of the corridor since the velocity profile 
should be symmetrically distributed with respect to the mid-axis of the 
corridor. Fig.~\ref{fig:appendix_2:1} shows the computed mean velocity for the 
bottom side profile as function of $\alpha$.\\

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\columnwidth]
{./fig_flux_model.eps}
\caption{\label{fig:appendix_2:2} Mean flux of the bottom half of the 
individuals vs. the pedestrian (global) density $\rho$ (see text for details). 
Both axis are dimensionless. The number of individuals across the 
corridor was set to $N=10$, and the contacting threshold was set to $\rho_0=1$. 
The ``packing coefficient'' was set to $c=1/\mathcal{K}$ (and thus, making the 
term $c\alpha$ independent of friction). The dashed line corresponds to the 
flux at the low density regime (say, $\langle v_i\rangle=1$).   }
\end{figure}

@@

We want to emphasize that the proposals stated in Refs.~\cite{parisi2,johansson,kwak}  only apply under normal conditions. If a crowd is under high levels of anxiety, pedestrians will neither keep distance between each other, nor will feel the urge to see an ``attraction". Thus, studying the friction coefficients may be a critical factor to properly reproduce the dynamics of a massive evacuation under stress.\\

@@

In this Appendix we show that the friction modification does not alter already studied behaviors of pedestrian dynamics (see Refs.~\cite{Dorso1,Dorso2,Dorso3}). In order to check this out, we performed numerical simulations of a room of 225 pedestrian escaping through a narrow door (bottleneck enclosure). The evacuation time (clearance time) $t_e$ is a function of the desired velocity $v_d$. This function has a minimum such that below it, $t_e$ is a decreasing function of $v_d$ while above, the tendency is reversed. It means that the evacuation process is optimum at moderated $v_d$. This is a well known phenomenon called the Faster-is-Slower effect. This effect was reported in numerical simulations \cite{Dorso1,Dorso5} and it has also experimental support \cite{parisi1}. \\

@@

Another pedestrian dynamics phenomenon is the lane formation reported in bidirectional flows \cite{feliciani1,helbing5,guo1,qiao1}. This means, individuals spontaneously organize in lanes of uniform walking direction, if the pedestrian density is high enough. One of the major achievements of the Social Force Model is its ability to reproduced this emergent phenomenon in which the microscopic interactions between pedestrians is enough to produce a global pattern. We run a bidirectional flow simulation (not shown) and verified that the friction increment still reproduces the lane formation process.\\

@@

Fig.~\ref{rate_fis} exhibits the ratio between the evacuation time for the friction modified model and the evacuation time for the original model. This ratio is shown as a function of the desired velocity. As the desired velocity increases, the rate of evacuation times approaches a constant value $\sim$ 9. This result is in agreement with the formula Eq.~(9) in Ref.~\cite{sticco}. The friction force dominates in the range of high desired velocities since it promotes the collisions between pedestrians. within this regime, the Eq. (9) of  Ref.~\cite{sticco} becomes a good approximation of the evacuation time. Thus, it is reasonable that dividing $t_e$ (for the enhanced friction) by $t_e$ (for the original model) yields a constant value for high $v_d$.

@@

@@

@@

@@

\bibitem{li1}
Li, M., Zhao, Y/., He, L., Chen, W. and Xu, X. 
2015. "The parameter calibration and optimization of social force model for the 
real-life 2013 Ya’an earthquake evacuation in China." Safety science 79, 
243-253.
{\path{https://doi.org/10.1016/j.ssci.2015.06.018}}

>>>>

\textcolor{red}{In the last decades, many microscopic models 
for crowd dynamics were developed. The class of force-based model was proposed by Hirai 
and Tarui (1975) introducing the class of \textit{force based} models 
\cite{Hirai}. These models assimilate the interaction between 
pedestrians and their environment using different kinds of forces inspired on physical systems. For 
example, Okazaki (1979) modeled the movement of pedestrians as 
a magnetized object immersed a magnetic in field \cite{Okazaki}.}\\

@@

\textcolor{red}{Our work focuses on data from Ref.~\cite{helbing3}. 
Further measurements can be found in 
Refs.~\cite{Boltes,lohner1,Dridi1,Dridi2,Baqui1}.
The PedFlow model developed by L\"ohner \textit{et al}. Ref.~\cite{lohner2}  
is an alternative to the Helbing model. Togashi \textit{et al}. improved the 
calibration of this model using a Kalman filter ensemble \cite{Togashi1}. }\\

The high density regime appears to be the most cumbersome one. Caution was 
claimed when (automatically) transferring the usual ``calibrated'' parameters 
of the SFM to this regime. It was argued that the pedestrians' body size 
distribution and the ``situational context'' are somewhat responsible for the 
unexpected departure from these parameters \cite{kwak,johansson1}. But other 
researchers pointed out that this departure actually expresses the lack of a 
mechanism to properly handle the pedestrians' ``required space to move''. Some 
modifications to the basic SFM were then proposed to overcome this difficulty 
\cite{parisi2,seyfried2}. \\

@@

Inspired from vehicular traffic dynamic studies, many researches on pedestrian 
dynamics focus their attention on the relation between the flow and the density 
of a moving crowd. This relation is represented by ``fundamental diagram" and it 
has become one of the most common ways to characterize the pedestrians' dynamics 
along a corridor in unidirectional and bidirectional flows 
\cite{seyfried1,fruin1,mori1,polus1,jelic1}. \\

@@

We implemented special modules in C++ for upgrading the LAMMPS 
capabilities to attain the Social Force Model simulations. We also checked over 
the LAMMPS output with previous computations (see Refs. \cite{Dorso2, 
Dorso1,Dorso3, Dorso4,Dorso6}).\\

@@

@@

Recall from Eq.~(\ref{granular}), that the friction force depends on the 
compression and the relative velocity among pedestrians. The compression levels 
remain the same in the three cases since the compression only depends on the 
density (which is fix at $\rho=6$ for the three color maps). Thus, the 
differences between Figs.~\ref{abswg_width10}, \ref{abswg_width15} and 
\ref{abswg_width22} can only explained by the increment of the relative velocity 
between individuals. \\

@@

A boundary condition needs to be imposed in order to solve 
Eq.~(\ref{eqn_appendix_2:3}) for $\alpha\neq 0$. We fix 
$v_i=v_{i+1}$ at the middle of the corridor since the velocity profile 
should be symmetrically distributed with respect to the mid-axis of the 
corridor. Fig.~\ref{fig:appendix_2:1} shows the computed mean velocity for the 
bottom side profile as function of $\alpha$.\\

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\columnwidth]
{./fig_flux_model.eps}
\caption{\label{fig:appendix_2:2} Mean flux of the bottom half of the 
individuals vs. the pedestrian (global) density $\rho$ (see text for details). 
Both axis are dimensionless. The number of individuals across the 
corridor was set to $N=10$ and the contacting threshold was set to $\rho_0=1$. 
The ``packing coefficient'' was set to $c=1/\mathcal{K}$ (and thus, making the 
term $c\alpha$ independent of friction). The dashed line corresponds to the 
flux at the low density regime (say, $\langle v_i\rangle=1$).   }
\end{figure}

@@

We want to emphasize that the proposals stated in 
Refs.~\cite{kwak,parisi2,johansson}  only apply under normal conditions. 
If a crowd is under high levels of anxiety, pedestrians will neither keep 
distance between each other, nor will feel the urge to see an ``attraction". 
Thus, studying the friction coefficients may be a critical factor to properly 
reproduce the dynamics of a massive evacuation under stress.\\

@@

In this Appendix we show that the friction modification does not alter already 
studied behaviors of pedestrian dynamics (see 
Refs.~\cite{Dorso2,Dorso1,Dorso3}). In order to check this out, we performed 
numerical simulations of a room of 225 pedestrian escaping through a narrow door 
(bottleneck enclosure). The evacuation time (clearance time) $t_e$ is a function 
of the desired velocity $v_d$. This function has a minimum such that below it, 
$t_e$ is a decreasing function of $v_d$ while above, the tendency is reversed. 
It means that the evacuation process is optimum at moderated $v_d$. This is a 
well known phenomenon called the Faster-is-Slower effect. This effect was 
reported in numerical simulations \cite{Dorso1,Dorso5} and it has also 
experimental support \cite{parisi1}. \\

@@

Another pedestrian dynamics phenomenon is the lane formation reported in 
bidirectional flows \cite{helbing5,feliciani1,guo1,qiao1}. This means, 
individuals spontaneously organize in lanes of uniform walking direction, if the 
pedestrian density is high enough. One of the major achievements of the Social 
Force Model is its ability to reproduced this emergent phenomenon in which the 
microscopic interactions between pedestrians is enough to produce a global 
pattern. We run a bidirectional flow simulation (not shown) and verified that 
the friction increment still reproduces the lane formation process.\\

@@

Fig.~\ref{rate_fis} exhibits the ratio between the evacuation time for the 
friction modified model and the evacuation time for the original model. This 
ratio is shown as a function of the desired velocity. As the desired velocity 
increases, the rate of evacuation times approaches a constant value $\sim$ 9. 
This result is in agreement with the formula Eq.~(9) in Ref.~\cite{sticco}. The 
friction force dominates in the range of high desired velocities since it 
promotes the collisions between pedestrians. within this regime, the Eq. (9) of  
Ref.~\cite{sticco} becomes a good approximation of the evacuation time. Thus, it 
is reasonable that dividing $t_e$ (for the enhanced friction) by $t_e$ (for the 
original model) yields a constant value for high $v_d$.  \\

@@

%37

@@

%43

@@

%35

@@

%40

\bibitem{li1}
Li, M., Zhao, Y., He, L., Chen, W. and Xu, X. 
2015. "The parameter calibration and optimization of social force model for the 
real-life 2013 Ya’an earthquake evacuation in China." Safety science 79, 
243-253.
{\path{https://doi.org/10.1016/j.ssci.2015.06.018}}
